# PLOTS MAPS




##  Sources with maps


1. [gadm](http://gadm.org/) - a database of global administrative boundaries
2. countries - polygons for all countries
3. SRTM - hole-filled CGIAR-SRTM (90 m resolution) Projections
4. alt - altitude (elevation) aggregated from SRTM 90 m resolution data between -60 and 60 latitude. 
5. worldclim - a database of global interpolated climate data

##  Administrative maps

###  p:ggplot2


Ploting only outline of the map.

```{r}
require(sp)
require(ggplot2)
require(maptools)

#loading to the workspace variable 'gadm' from web site
load(url("http://biogeo.ucdavis.edu/data/gadm2/R/POL_adm1.RData"))
wojewodztwa <- gadm
load(url("http://biogeo.ucdavis.edu/data/gadm2/R/POL_adm2.RData"))
powiaty <- gadm


#first way
ggplot(wojewodztwa, aes(x = long, y = lat, group=group)) + geom_path()
ggplot(powiaty, aes(x = long, y = lat, group=group)) + geom_path()



#second way with ggmap function
woj_fort <- fortify(wojewodztwa)
ggplot(data=woj_fort, aes(map_id=id, x=long, y=lat, group=group)) + geom_map(map = woj_fort, fill='yellow') + geom_path()

```



Downloading shp files<br>

Trzeba sciagnac plik z mapa, np. ze strony \href{http://gadm.org/country}{(gadm)}. W sciagnietym pliku mapy beda w roznych formatach. Jest rowniez potrzebny ponizej format shp. Mapa Polski jest dostepna w trzech wersjach:
1. kontury samej polski
2. podzial na wojewodztwa
3. podzial na powiaty
\end{enumerate}
**Uwaga**. Zeby dobrze dzialalo nie wystarczy sam plik shp. Lepiej aby w katalogy byly pozostale pliki. Widocznie sa tez wykorzystywane.

```{r, opts.label='k'}
#polska z wojewodztwami
require(maptools)
polska_shp <- readShapePoly("C:/Users/ppp/Desktop/BIBLIA_R/MANUAL/2_WIZUALIZACJA/1_PODSTAWOWE_PAKIETY/demos/mapy/POL_adm1.shp")
polska_map <- fortify(polska_shp)
ggplot(polska_map, aes(x = long, y = lat, group=group)) + geom_path()
```


<br> Manual map: example from help
```{r, o.l='ko', s.after=TRUE}
ids <- factor(c("1.1", "2.1", "1.2", "2.2", "1.3", "2.3"))

values <- data.frame(
  id = ids,
  value = c(3, 3.1, 3.1, 3.2, 3.15, 3.5)
)

positions <- data.frame(
  id = rep(ids, each = 4),
  x = c(2, 1, 1.1, 2.2, 1, 0, 0.3, 1.1, 2.2, 1.1, 1.2, 2.5, 1.1, 0.3,
  0.5, 1.2, 2.5, 1.2, 1.3, 2.7, 1.2, 0.5, 0.6, 1.3),
  y = c(-0.5, 0, 1, 0.5, 0, 0.5, 1.5, 1, 0.5, 1, 2.1, 1.7, 1, 1.5,
  2.2, 2.1, 1.7, 2.1, 3.2, 2.8, 2.1, 2.2, 3.3, 3.2)
)

ggplot(values) + geom_map(aes(map_id = id), map = positions) +
  expand_limits(positions)
ggplot(values, aes(fill = value)) +
  geom_map(aes(map_id = id), map = positions) +
  expand_limits(positions)
ggplot(values, aes(fill = value)) +
  geom_map(aes(map_id = id), map = positions) +
  expand_limits(positions) + ylim(0, 3)

```




<br> Extracting detaild information
```{r, o.l='ko', s.before=TRUE}

powiaty@data$ID_0
powiaty@data$NAME_0
powiaty@data$ID_1
powiaty@data$NAME_1
powiaty@data$ID_2
powiaty@data$NAME_2
powiaty@data$VARNAME_2


wojewodztwa@data$ID_0
wojewodztwa@data$NAME_0
wojewodztwa@data$ID_1
wojewodztwa@data$NAME_1
wojewodztwa@data$VARNAME_1

```




<br> Named elements
```{r}
library(raster)
library(rgdal)
library(rgeos)
library(ggplot2)
library(dplyr)
require(sp)


load(url("http://biogeo.ucdavis.edu/data/gadm2/R/POL_adm1.RData"))
wojewodztwa <- gadm

polska_woj <- fortify(wojewodztwa)
polska_woj$id <- as.integer(polska_woj$id)
dat <- data.frame(id=1:(length(wojewodztwa@data$NAME_1)), wojewodztwa=wojewodztwa@data$VARNAME_1)

polska_woj_df <- merge(polska_woj, dat, by='id')

centers <- data.frame(gCentroid(wojewodztwa, byid=TRUE))
centers$wojewodztwa <- dat$wojewodztwa

ggplot() + 
  geom_map(data=polska_woj_df, map=polska_woj_df, aes(map_id=id, x=long, y=lat, group=group), color='#ffffff', fill='#bbbbbb', size=0.25) + 
  geom_text(data=centers, aes(label=wojewodztwa, x=x, y=y), size=3) + 
  coord_map() + 
  labs(x='', y='', title='Poland')

```



<br> Extracting subelements example 1

```{r}

require(sp)
require(ggplot2)
load(url("http://biogeo.ucdavis.edu/data/gadm2/R/POL_adm1.RData"))
wojewodztwa <- gadm

polska_woj<- fortify(wojewodztwa)
polska_slaskie <- polska_woj[polska_woj[,'id']==13,]
ggplot(polska_slaskie, aes(x=long, y=lat, group=group)) + geom_path()

```



<br> Extracting subelements example 2

```{r}
require(sp)
require(ggplot2)
load(url("http://biogeo.ucdavis.edu/data/gadm2/R/POL_adm2.RData"))
powiaty <- gadm

id_powiaty_slaskie <- powiaty@data %>% filter(ID_1==13) %>% dplyr::select(ID_2)

polska_pow <- fortify(powiaty)

powiaty_slaskie <- polska_pow %>% filter(id %in% as.character(unlist(id_powiaty_slaskie)))
ggplot(powiaty_slaskie, aes(x=long, y=lat, group=group)) + geom_path()

```



<br> Maps with colors: example 1

```{r}
require(sp)
require(ggplot2)
load(url("http://biogeo.ucdavis.edu/data/gadm2/R/POL_adm1.RData"))
wojewodztwa <- gadm
polska_woj <- fortify(wojewodztwa)

unemployment_data <- data.frame(id=rownames(wojewodztwa@data), unemployment=runif(16,min = 5,max = 12))

polska_woj <- merge(polska_woj, unemployment_data, by='id')

ggplot(polska_woj, aes(x=long, y=lat, group=group)) + 
  geom_polygon(aes(fill=unemployment)) + 
  geom_path(color='grey') +
  scale_fill_gradient(low='light green', high='dark green') +
  coord_map()


```


Maps with colors: example 2 with faceting

```{r}
require(sp)
require(ggplot2)
load(url("http://biogeo.ucdavis.edu/data/gadm2/R/POL_adm1.RData"))
wojewodztwa <- gadm
polska_woj <- fortify(wojewodztwa)

unemployment_data <- data.frame(id=rep(rownames(wojewodztwa@data),2), unemployment=runif(32,min = 5,max = 12), year=rep(c(2020,2021), each=16) )

polska_woj <- merge(polska_woj, unemployment_data, by='id')

ggplot(polska_woj, aes(x=long, y=lat, group=group)) + 
  geom_polygon(aes(fill=unemployment)) + 
  geom_path(color='grey') +
  scale_fill_gradient(low='light green', high='dark green') +
  coord_map() + facet_grid(.~year)

```




<br>Maps with colors: example 3 descrete categories

```{r, o.l='k', s.before=TRUE, s.after=TRUE}

polska_woj <- fortify(wojewodztwa)

unemployment_continuous <- runif(32,min = 4,max = 12)
unemployment_discrete <- cut(unemployment_continuous, 
                             breaks=c(4,6,8,10,12), 
                             include.lowest=TRUE, 
                             ordered_results=TRUE)


require(plotrix)
colors <- gray.colors(nlevels(unemployment_discrete), start = 0.9, end = .1)

require(plyr)
unemployment_colors <- as.character(mapvalues(unemployment_discrete, levels(unemployment_discrete), colors))

unemployment_data <- data.frame(id=rep(rownames(wojewodztwa@data),2), unemployment=unemployment_discrete, year=rep(c(2020,2021), each=16) )

polska_woj_1 <- merge(polska_woj, unemployment_data, by='id')

ggplot(polska_woj_1, aes(x=long, y=lat, group=group)) + 
  geom_polygon(aes(fill=unemployment)) + 
  geom_path(color='black') +
  scale_fill_manual(values = colors) +
  coord_map() + facet_grid(.~year)

```



##  p:leaflet


###  Interactive map in shiny with wojewodztwa


```{r}

require(shiny)
require(l.shiny)
require(ggplot2)
require(DT)
require(rhandsontable)
require(leaflet)

u <- shinyUI(fluidPage(
	
	leafletOutput('map1'),
	textOutput('wojewodztwo')
	
))


s <- function(input, output, session){
	
output$map1 <- renderLeaflet({
	
	gadm <- readRDS('C:/Users/ppp/Desktop/Temporary_files/POL_adm1.rds')

	pal <- colorNumeric(
  palette = "Blues",
  domain = 1:16
)
	gadm <- readRDS('C:/Users/ppp/Desktop/Temporary_files/POL_adm1.rds')
	
	leaflet() %>% addTiles() %>% addPolygons(data=gadm, weight=2, fillColor = ~pal(1:16) , layerId = paste0("marker", 1:16) )
# popup = ~pal(1:16), 
})
	

output$wojewodztwo <- renderText({
	z <- input$map1_shape_click$id
	print(unlist(z))
})

}

shinyApp(ui=u, server=s)


```

###  removing elements from map

```{r}

require(leaflet)


library(shiny)

ui <- fluidPage(leafletOutput("map1"))

server <- function(input, output, session) {
    output$map1 <- renderLeaflet({
        leaflet() %>% addCircleMarkers(lng = runif(10), lat = runif(10), 
            layerId = paste0("marker", 1:10))
    })
    
    observeEvent(input$map1_marker_click, {
        leafletProxy("map1", session) %>% removeMarker(input$map1_marker_click$id)
    })
}

shinyApp(ui, server)


```


##  physical maps

##  google static maps



```{r}

require(ggplot2)
library(ggmap)
al1 <- get_map(location = c(lon = -86.304474, lat = 32.362563), 
  zoom = 11, 
  maptype = 'terrain') #other options: "terrain", "satellite", "roadmap", "hybrid", "toner", "watercolor"

w <- gglocator(n = 5, message = FALSE, xexpand = c(0.05, 0), yexpand = c(0.05, 0))

ggmap(al1)+ geom_point(data=w, aes(x=lon, y=lat), size=8)


#faceting

w <- cbind(w, id=c(1,1,2,2,2))
ggmap(al1)+ geom_point(data=w, aes(x=lon, y=lat), size=8) + facet_grid(.~id)

```


##  google dynamic maps


##  p:Leaflet

```{r}

devtools::install_github("rstudio/leaflet")

library(leaflet)
m = leaflet() %>% addTiles()
m  # a map with the default OSM tile layer


#map with zoom
m = m %>% setView(-93.65, 42.0285, zoom = 17)
m


#adding popup
m %>% addPopups(-93.65, 42.0285, 'Here is the <b>Department of Statistics</b>, ISU')

#adding circle with popuu
m %>% addCircles(-93.65, 42.0285, popup = 'Here is the <b>Department of Statistics</b>, ISU')



#adding points

set.seed(123)
m = leaflet() %>% addTiles()
rand_lng = function(n = 10) rnorm(n, -93.65, .01)
rand_lat = function(n = 10) rnorm(n, 42.0285, .01)

# circles (units in metres)
m %>% addCircles(rand_lng(50), rand_lat(50), radius = runif(50, 10, 200))


```

